A Comparison of Dimensionality Reduction Techniques for Robotic Motion Planning

### 2016-05-14 paper/paper.tex
finished conclusion

        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
        \STATE Generate random obstacles based off of obstacle information
        \STATE Fit dimensionality reduction transform to the space
        \STATE Transform the space
        \STATE Run PRM* or RRTConnect over the space
        \STATE Inverse transform the path back to the original dimension
        \STATE Check for collisions
        \RETURN Path\_length, duration, num\_failed
    \end{algorithmic}
\end{algorithm}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM* and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we find to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found. Note, that the length of the path is computed after
it has been transformed back into the original high dimensional space in order
to make a fair comparison. A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

@@

Fig.~\ref{fig:plots} shows plots of various measures of performance using
different methods of dimensionality reduction for motion planing. Please note
that the Y-axes of these plots are not equivalent in the left and right column.
The left and right columns show the results for RRTConnect and PRM algorithms
respectively.  The first, second, and third row show the average path length,
computational time in seconds, and number of failed runs out of 100
respectively. For each dimensionality reduction technique, there are three bars
presenting the results for 250, 300, and 350 randomly generated obstacles
respectively. From the data we can see that both planners produced similarly
sized paths for each dimensionality reduction technique, and that by not
reducing the dimension, the planner is able to find a path roughly two times
shorter.

@@

In terms of computational time, there is no significant difference between
Truncated SVD, PCA, Randomized PCA, planning in the original space when
planning with the RRTConnect algorithm. Whereas using Feature Agglomeration
took slightly longer to compute for each set number of obstacles. When using
PRM, the computational time was significantly improved by reducing the
dimension for all methods.  Based on the data, using PRM without dimensionality
reduction took roughly twice as long to compute for each method. This differing
result between PRM and RRTConnect is likely due to the fact that PRM is a
multi-query planning algorithm meaning that it constructs a data-structure that
can be used for multiple planning queries whereas RRTConnect is only looking to
find a single path. This means that the PRM algorithm tries to fully represent
the connectivity of the environment whilst RRTConnect needs to determine the
connectivity directed towards a goal. The reason dimensionality reduction
significantly improves the computational time for PRM is that it is much faster
to exhaust the search space at lower dimensions as opposed to higher
dimensions.

The RRTConnect algorithm had a much lower failure rate than PRM in all
respective dimensionality reduction technique. This is due to the fact that
RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
connecting points to a growing graph. For both algorithms, a similar trend
exists for failure rates when planning in the reduced dimension space. For both
planners, Feature Agglomeration failed more frequently than the rest of the
techniques, and planning in the high dimensional space had the smallest failure
rate. The higher failure rate for Feature Agglomeration is because Feature
Agglomeration works best when features from the data can be combined due to
repetition and redundancy. Since this data was in the $\mathbb{R}^n$ and the
random obstacles were generated all over the space, Feature Agglomeration
failed to produce an embedding that also worked for planning in the lower
dimensional space. Aside from Feature Agglomeration, the dimensionality
reduction techniques failed to find a path to the goal configuration between 20
and 60 percent of the time for PRM and between 5 and 35 percent for RRTConnect.
This is due to the projection of the higher dimensional obstacles into a lower
dimension causing occlusions in the lower dimensional space where they would
not exist in the higher dimension. However, these dimensionality reduction
techniques can be seen as independent random variables for each run, therefore,
you would be able to run the algorithm again with the same probability of
failure. As the number of obstacles increased, so did the failure rate because
the lower dimensional spaces become more densely populated with obstacles
causing more occlusions.

From the data, we can see that dimensionality reduction techniques can reduce
the optimality of the planning algorithm, but in multi-query planning
algorithms, these techniques can significantly reduce the time needed to
compute a solution. Based on the data, it seems that the Truncated SVD
algorithm can be used to produce the lowest number of failures and produces
similar results in the path length and computation time  as compared to the
other techniques used. Also, we Can see that Feature Agglomeration fails to
find a solution path more frequently than the other techniques, but produces
shorter paths when it does find a path.

@@

>>>>

        \STATE Specify dimension, dimensionality reduction technique, and obstacle information
        \STATE Generate random obstacles based off of obstacle information
        \STATE Fit dimensionality reduction transform to the space
        \STATE Transform the space
        \STATE Run PRM or RRTConnect over the space
        \STATE Inverse transform the path back to the original dimension
        \STATE Check for collisions
        \RETURN Path\_length, duration, num\_failed
    \end{algorithmic}
\end{algorithm}

After generating random obstacles, we ran Feature Agglomeration, Truncated SVD,
PCA, and Randomized PCA to reduce the dimensionality of the space by using the
sampled points within each $n$-sphere as training data. We then ran PRM and
RRTConnect on each of these reduced dimensionality spaces, as well as on the
fully-dimensional space as a control. We then ran an inverse transform on the
paths that we found to return them to the fully-dimensional space in order to
verify that the path in the reduced dimension did not come into collision with
obstacles. We collected data on the number of runs that fail to find a
solution, the time it takes to run the algorithm, and the length of the
solution if one was found. Note that the length of the path is computed after
it has been transformed back into the original high dimensional space in order
to make a fair comparison. A summary of this paragraph can be found in
Algo.~\ref{alg:DR}.

@@

Fig.~\ref{fig:plots} shows plots of various measures of performance using
different methods of dimensionality reduction for motion planing. Please note
that the Y-axes of these plots are not equivalent in the left and right column.
The left and right columns show the results for the RRTConnect and PRM algorithms
respectively.  The first, second, and third row show the average path length,
computational time in seconds, and number of failed runs out of 100
respectively. For each dimensionality reduction technique, there are three bars
presenting the results for 250, 300, and 350 randomly generated obstacles
respectively. From the data we can see that both planners produced similarly
sized paths for each dimensionality reduction technique, and that by not
reducing the dimension, the planner is able to find a path roughly two times
shorter.

@@

In terms of computational time, there is no significant difference between
Truncated SVD, PCA, Randomized PCA, and planning in the original space when
planning with the RRTConnect algorithm, whereas using Feature Agglomeration
took slightly longer to compute for each set number of obstacles. When using
PRM, the computational time was significantly improved by reducing the
dimension for all methods.  Based on the data, using PRM without dimensionality
reduction took roughly twice as long to compute for each method. This differing
result between PRM and RRTConnect is likely due to the fact that PRM is a
multi-query planning algorithm, meaning that it constructs a data-structure that
can be used for multiple planning queries, whereas RRTConnect is only looking to
find a single path. This means that the PRM algorithm tries to fully represent
the connectivity of the environment whilst RRTConnect needs to determine the
connectivity directed towards a goal. The reason dimensionality reduction
significantly improves the computational time for PRM is that it is much faster
to exhaust the search space at lower dimensions as opposed to higher
dimensions.

The RRTConnect algorithm had a much lower failure rate than PRM in all
respective dimensionality reduction technique. This is due to the fact that
RRTConnect is growing a tree towards the goal and PRM is randomly sampling and
connecting points to a growing graph. For both algorithms, a similar trend
exists for failure rates when planning in the reduced dimension space. For both
planners, Feature Agglomeration failed more frequently than the rest of the
techniques, and planning in the high dimensional space had the smallest failure
rate. The higher failure rate for Feature Agglomeration is because Feature
Agglomeration works best when features from the data can be combined due to
repetition and redundancy. Since this data was in $\mathbb{R}^n$ and the
random obstacles were generated all over the space, Feature Agglomeration
failed to produce an embedding that also worked for planning in the lower
dimensional space. Aside from Feature Agglomeration, the dimensionality
reduction techniques failed to find a path to the goal configuration between 20
and 60 percent of the time for PRM and between 5 and 35 percent for RRTConnect.
This is due to the projection of the higher dimensional obstacles into a lower
dimension causing occlusions in the lower dimensional space where they would
not exist in the higher dimension. However, these dimensionality reduction
techniques can be seen as independent random variables for each run; therefore,
you would be able to run the algorithm again with the same probability of
failure. As the number of obstacles increased, so did the failure rate because
the lower dimensional spaces become more densely populated with obstacles,
causing more occlusions.

From the data, we can see that dimensionality reduction techniques can reduce
the optimality of the planning algorithm, but in multi-query planning
algorithms, these techniques can significantly reduce the time needed to
compute a solution. Based on the data, it seems that the Truncated SVD
algorithm can be used to produce the lowest number of failures and produces
similar results in the path length and computation time  as compared to the
other techniques used. Also, we can see that Feature Agglomeration fails to
find a solution path more frequently than the other techniques, but produces
shorter paths when it does find a path.

@@

Thus, in this project we compared several dimensionality reduction techniques against each other and against the option of no reduction. Our experiments showed that, at least for the case of random obstacles in $\mathbb{R}^n$, Feature Agglomeration performs very poorly, while Truncated SVD slightly outperforms PCA and Randomized PCA. We also found that, for multi-query planning algorithms, dimensionality reduction significantly reduced the calculation time. However, this benefit was not seen for the single-query RRTConnect. Unfortunately, dimensionality reduction seems to come at a cost - in our tests, dimensionality reduction doubled path length and often caused the planning algorithm to fail. In conclusion, in our tests, dimensionality reduction improved the computation time of the multi-query planning algorithm but resulted in longer path lengths and more failures than the nonreduced-dimension case.
